### 2023-04-10 2023-1-19/efficient_way_for_counting.tex
atva2023 first version

\documentclass[runningheads]{llncs}

\usepackage{standalone}
\usepackage{import}
\usepackage{subcaption}
\usepackage{pgfplots}
\usepackage{color}
\usepackage{tikz}
\usepackage{pgfplots}
\usepgfplotslibrary{fillbetween}
\pgfplotsset{compat=1.16}
\usetikzlibrary {arrows.meta, automata, positioning, fit}
\usepackage{url}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage[]{float}
% set algorithmicx
\renewcommand{\algorithmicrequire}{\textbf{Input:}}  % Use Input in the format of Algorithm  
\renewcommand{\algorithmicensure}{\textbf{Output:}} % Use Output in the format of Algorithm

\newcommand*{\red}[1]{\textcolor{red}{#1}}

\newcommand*{\myvec}[1]{\vec{#1}}
\newcommand*{\regex}{\mathcal{R}}
\newcommand*{\strsort}{\verb|Str|}
\newcommand*{\intsort}{\verb|Int|}
\newcommand*{\lan}{\mathcal{L}}
\newcommand*{\highlight}[1]{\textbf{\textit{#1}}}
\newcommand*{\aut}{\mathcal{A}}
\newcommand*{\algfun}[1]{\texttt{#1}}
\newcommand*{\strlen}[1]{\texttt{len}(#1)}
\newcommand*{\myemph}[1]{\textcolor{red}{#1}}
\newcommand*{\mult}[1]{$\times$}
%%
%% end of the preamble, start of the body of the document source.
\begin{document}
%%
%% The "title" command has an optional parameter,
%% allowing the author to define a "short title" to be used in page headers.
\title{Efficient Methods for Solving String Constraints with Bounded Repetition}

%%
\author{Denghang Hu\inst{1} \and
  Zhilin Wu\inst{1}}
\authorrunning{D.Hu et al.}
% \authornote{Both authors contributed equally to this research.}

\institute{State Key Laboratory of Computer Science,
  Institute of Software, Chinese Academy of Sciences, China\\
  \email{\{hudh,wuzl\}@ios.ac.cn}}

\maketitle

%% The abstract is a summary of the work to be presented in the
%% article.
\begin{abstract}
\subimport{sections/}{abstract.tex}
\end{abstract}

\section{Introduction} \label{sec:intro}
\subimport{sections/}{introduction.tex}

\section{Preliminaries} \label{sec:pre}
\subimport{sections/}{preliminaries.tex}

\section{Overview} \label{sec:overall}
\subimport{sections/}{overview.tex}

\section{Automaton Model} \label{sec:overview}
\subimport{sections/}{automaton_model.tex}

\section{Solve ESL Formula with Large Repetition Bound} \label{sec:algorithm}
\subimport{sections/}{algorithms.tex}

\section{Implementation and Experimental Results} \label{sec:implementation}
\subimport{sections/}{implementation.tex}

\section{Conclusion and Future Work} \label{sec:conclu}
\subimport{sections/}{conclusion.tex}

\end{document}

%%
%% End of file `sample-sigconf.tex'.