### 2023-04-11 2023-1-19/figures/cactus_plot_all.tex
atva2023 first version

\documentclass[11pt]{standalone}
\usepackage{color}
\usepackage{tikz}
\usepackage{pgfplots}
\usepgfplotslibrary{fillbetween}
\pgfplotsset{compat=1.16}
\begin{document}
\definecolor{colourCVC5}{HTML}{25333D}
\definecolor{colourOstrichCEA}{HTML}{0065AB}
\definecolor{colourOstrich}{HTML}{8939AD}
\definecolor{colourZ3str3}{HTML}{007E7A}
\definecolor{colourZ3seq}{HTML}{CD3517}
\definecolor{colourZ3str3RE}{HTML}{318700}
\definecolor{colourZ3-Trau}{HTML}{80746D}
\resizebox{.95\textwidth}{!}{\pgfplotsset{scaled x ticks=false}\pgfplotsset{scaled y ticks=false}\begin{tikzpicture}\begin{axis}[title=Total,xlabel=Solved instances,ylabel=Time (s),,legend columns=2,legend style={nodes={scale=0.5, transform shape}, fill=none,anchor=east,align=center },axis line style={draw=none}, xtick pos=left, ytick pos=left, ymajorgrids=true, legend style={draw=none},x post scale=2,y post scale=1]
\addplot[name path=pathCVC5 , colourCVC5, line width=1.5pt] coordinates {(1,0.004)(225,0.9749977728285077)(674,3.1805545657015593)(1123,5.703)(1572,8.438265033407571)(2021,11.531574610244988)(2470,15.075634743875279)(2919,19.097318485523385)(3368,23.513942093541203)(3817,28.284)(4266,33.33244543429844)(4715,38.708721603563475)(5164,44.401)(5613,50.33589755011136)(6062,56.600924276169266)(6511,63.173)(6960,69.98948106904231)(7409,77.14595100222718)(7858,84.631)(8307,92.40952561247217)(8756,100.53519599109131)(9205,109.03022717149221)(9654,117.91206904231626)(10103,127.1663407572383)(10552,136.837)(11001,146.92912249443208)(11450,157.46868596881959)(11899,168.51312026726058)(12348,180.07232962138085)(12797,192.1541959910913)(13246,204.78042093541202)(13695,218.04313363028953)(14144,232.00983073496658)(14593,246.74662138084634)(15042,262.27243207126946)(15491,278.68448552338526)(15940,296.06152338530063)(16389,314.4960133630289)(16838,333.9485055679287)(17287,354.4259688195991)(17736,375.96304899777283)(18185,398.6540556792873)(18634,422.5968173719376)(19083,447.8012538975501)(19532,474.371004454343)(19981,502.39159242761696)(20430,531.845280623608)(20879,562.9128819599109)(21328,595.7690712694878)(21777,630.5182316258351)(22226,667.4171893095769)(22675,706.6891113585746)(23124,748.524723830735)(23573,792.9852828507795)(24022,840.2546636971047)(24471,890.3767973273942)(24920,943.5922204899778)(25369,1000.1014365256125)(25818,1060.165797327394)(26267,1123.928570155902)(26716,1191.6269866369712)(27165,1263.5346726057908)(27614,1339.8836258351894)(28063,1421.4921870824055)(28512,1509.265877505568)(28961,1603.7898262806236)(29410,1705.7232071269486)(29859,1815.975158129176)(30308,1935.8308218262805)(30757,2066.3378129175944)(31206,2207.6831870824053)(31655,2360.5879977728287)(32104,2524.3051202672605)(32553,2699.568131403118)(33002,2887.1032249443206)(33451,3087.4128195991093)(33900,3301.4712672605788)(34349,3529.438995545657)(34798,3772.0046904231626)(35247,4029.107385300668)(35696,4301.784002227171)(36145,4592.399788418708)(36594,4903.131832962137)(37043,5237.107131403119)(37492,5598.406133630289)(37941,5992.969233853007)(38390,6426.496371937639)(38839,6909.088826280624)(39288,7452.511380846326)(39737,8072.1869844098)(40186,8790.347349665924)(40635,9647.260521158129)(41084,10728.830126948777)(41533,12233.169844097996)(41982,14371.070224944322)(42431,17689.691844097993)(42880,23713.308975501113)(43329,34464.926198218265)(43776.5,51319.2588161435)(44001,62564.102)};
\addlegendentry{CVC5}

\addplot[name path=pathOstrichCEA , colourOstrichCEA, line width=1.5pt] coordinates {(1,1.001)(251,297.57713173652695)(752,919.7552734530939)(1253,1557.5426886227544)(1754,2204.4396926147706)(2255,2859.376377245509)(2756,3521.8273872255486)(3257,4190.722225548902)(3758,4865.928748502994)(4259,5547.6787385229545)(4760,6236.603149700599)(5261,6933.963383233533)(5762,7640.762471057885)(6263,8358.542546906187)(6764,9088.009331337325)(7265,9829.516295409183)(7766,10584.155027944113)(8267,11351.854083832335)(8768,12132.038473053894)(9269,12922.84225748503)(9770,13722.257604790419)(10271,14529.294508982037)(10772,15342.857842315369)(11273,16162.547750499003)(11774,16987.797055888226)(12275,17818.525373253495)(12776,18654.640872255488)(13277,19495.92945708583)(13778,20342.265848303392)(14279,21193.5768003992)(14780,22049.840229540918)(15281,22911.280876247507)(15782,23777.89823752495)(16283,24649.420652694607)(16784,25525.914844311377)(17285,26407.47073253493)(17786,27293.949608782437)(18287,28185.26401996008)(18788,29081.597538922157)(19289,29982.794083832334)(19790,30888.764532934132)(20291,31799.552532934133)(20792,32714.965686626747)(21293,33634.779389221556)(21794,34558.914994011975)(22295,35487.432938123755)(22796,36420.41224550898)(23297,37357.921341317364)(23798,38300.12236526946)(24299,39247.05026946108)(24800,40198.6993253493)(25301,41155.420005988024)(25802,42117.09206986028)(26303,43083.934487025945)(26804,44055.638161676645)(27305,45032.20359081836)(27806,46014.43762275449)(28307,47002.47610978044)(28808,47996.60786626746)(29309,48996.72237325349)(29810,50002.71481636727)(30311,51014.77076447106)(30812,52033.40894211577)(31313,53058.83036726547)(31814,54091.27846506986)(32315,55130.93350698603)(32816,56178.196696606785)(33317,57233.01951297405)(33818,58295.94183832336)(34319,59367.47835329341)(34820,60448.25150898203)(35321,61539.139007984035)(35822,62640.02783033932)(36323,63751.15949301397)(36824,64873.163516966066)(37325,66006.96277045908)(37826,67153.20774251496)(38327,68312.41110578843)(38828,69485.79568263472)(39329,70674.74570259481)(39830,71879.40951097805)(40331,73099.98904191617)(40832,74339.00549301397)(41333,75598.06305988024)(41834,76879.22272055889)(42335,78185.40860279441)(42836,79517.57918962077)(43337,80881.22110778444)(43838,82281.94577045909)(44339,83721.56675449101)(44840,85208.95984231538)(45341,86753.0387285429)(45842,88367.2720998004)(46343,90074.41964870259)(46844,91896.8463233533)(47345,93861.7311497006)(47846,96027.48211576846)(48347,98576.50299800398)(48848,102635.06975449102)(49112,108033.23818518518)(49127,108809.673)};
\addlegendentry{OstrichCEA}

\addplot[name path=pathOstrich , colourOstrich, line width=1.5pt] coordinates {(1,0.966)(242,253.6938115942029)(725,797.780635610766)(1208,1375.6963540372672)(1691,1971.2173726708074)(2174,2579.9891242236026)(2657,3200.1552774327124)(3140,3831.5469544513458)(3623,4473.097327122153)(4106,5124.210623188406)(4589,5784.478358178054)(5072,6452.883861283644)(5555,7128.7226770186335)(6038,7811.80366252588)(6521,8501.433523809525)(7004,9197.051412008283)(7487,9898.351950310558)(7970,10605.06412215321)(8453,11317.27030020704)(8936,12035.019962732918)(9419,12758.173621118012)(9902,13486.76760041408)(10385,14220.709316770186)(10868,14959.795536231884)(11351,15703.822519668738)(11834,16453.023033126294)(12317,17207.60976190476)(12800,17967.250432712215)(13283,18732.071757763973)(13766,19501.90405383023)(14249,20276.887979296065)(14732,21056.855813664595)(15215,21842.13654037267)(15698,22632.851068322983)(16181,23429.176252587993)(16664,24230.890689440996)(17147,25038.21724637681)(17630,25851.334213250517)(18113,26669.63826293996)(18596,27493.383886128366)(19079,28323.0271242236)(19562,29158.7289689441)(20045,30000.663378881985)(20528,30848.824979296067)(21011,31703.274051759836)(21494,32564.10966252588)(21977,33431.79361697723)(22460,34306.333964803314)(22943,35188.23764182194)(23426,36077.66052587992)(23909,36975.0900331263)(24392,37880.26252380952)(24875,38793.40235403727)(25358,39715.36438095238)(25841,40646.487689441)(26324,41586.495204968945)(26807,42535.87147412008)(27290,43494.791265010346)(27773,44463.5207184265)(28256,45442.81579710145)(28739,46433.07173706004)(29222,47434.819573498964)(29705,48449.619180124224)(30188,49478.66922360248)(30671,50522.00113043478)(31154,51581.44218426501)(31637,52657.74352795031)(32120,53751.45314699793)(32603,54863.35853830228)(33086,55994.589289855074)(33569,57145.683132505175)(34052,58318.82643478261)(34535,59515.03370600414)(35018,60737.58520703933)(35501,61988.47214492754)(35984,63269.9586563147)(36467,64585.57833333334)(36950,65939.52480745342)(37433,67336.00833747412)(37916,68780.93298964803)(38399,70278.50410351966)(38882,71836.31644513458)(39365,73463.24823602484)(39848,75170.24680331264)(40331,76972.25751759835)(40814,78886.19549068323)(41297,80940.91662111801)(41780,83159.28797722567)(42263,85568.35025051759)(42746,88213.17140786748)(43229,91128.96203933746)(43712,94354.33373498965)(44195,97990.84464389233)(44678,102163.55392546584)(45161,107056.2527805383)(45644,112920.6944989648)(46127,120179.79210559007)(46610,129819.74801035198)(47093,144581.69753830228)(47351.5,156224.77838235296)(47370,157248.051)};
\addlegendentry{Ostrich}

\addplot[name path=pathZ3str3 , colourZ3str3, line width=1.5pt] coordinates {(1,0.011)(186,2.2441590296495955)(557,6.933)(928,11.756)(1299,16.579323450134773)(1670,21.640361185983828)(2041,27.149)(2412,32.7522641509434)(2783,38.633)(3154,44.569)(3525,50.505)(3896,56.501285714285714)(4267,62.774)(4638,69.081)(5009,75.52557412398923)(5380,82.2)(5751,88.973)(6122,96.007)(6493,103.1635552560647)(6864,110.573)(7235,117.993)(7606,125.53147978436657)(7977,133.315)(8348,141.106)(8719,148.99271698113208)(9090,157.14)(9461,165.30301886792452)(9832,173.677)(10203,182.21)(10574,190.76738005390837)(10945,199.596)(11316,208.5)(11687,217.41128032345014)(12058,226.567)(12429,235.842)(12800,245.117)(13171,254.577)(13542,264.223)(13913,273.8815498652291)(14284,283.797)(14655,293.8208894878706)(15026,304.088)(15397,314.48077088948787)(15768,325.109)(16139,335.8730970350404)(16510,346.874)(16881,358.04226415094337)(17252,369.488)(17623,381.06956603773585)(17994,392.92)(18365,404.96041239892185)(18736,417.2412641509434)(19107,429.8)(19478,442.5950269541779)(19849,455.633102425876)(20220,468.952436657682)(20591,482.541)(20962,496.35875471698114)(21333,510.47258760107815)(21704,524.8864797843665)(22075,539.608)(22446,554.6164123989217)(22817,569.9552641509433)(23188,585.6441725067385)(23559,601.7121401617251)(23930,618.1685202156334)(24301,635.0817951482479)(24672,652.4683611859839)(25043,670.3729433962263)(25414,688.8481024258759)(25785,707.9234986522912)(26156,727.6508382749325)(26527,748.0661428571428)(26898,769.1473557951483)(27269,790.9199433962264)(27640,813.4528598382749)(28011,836.8241024258759)(28382,861.080665768194)(28753,886.3027628032345)(29124,912.5774959568732)(29495,939.9839460916443)(29866,968.653962264151)(30237,998.7923530997305)(30608,1030.6136603773584)(30979,1064.2623396226415)(31350,1100.220989218329)(31721,1138.6942614555257)(32092,1179.9984878706198)(32463,1224.7958005390835)(32834,1274.0089541778977)(33205,1328.8225849056605)(33576,1391.1415417789756)(33947,1463.6065067385443)(34318,1550.8098005390837)(34689,1661.4692075471698)(35060,1824.3313234501347)(35431,2141.3325256064686)(35802,2940.3429811320757)(36151.5,6899.040887195121)(36317,12695.69)};
\addlegendentry{Z3str3}

\addplot[name path=pathZ3seq , colourZ3seq, line width=1.5pt] coordinates {(1,0.011)(239.5,3.003002092050209)(717.5,9.1975)(1195.5,15.473012552301256)(1673.5,22.107)(2151.5,28.998301255230125)(2629.5,36.189257322175735)(3107.5,43.723)(3585.5,51.41517364016736)(4063.5,59.4635)(4541.5,67.71654184100419)(5019.5,76.30319037656903)(5497.5,85.1595)(5975.5,94.25958786610879)(6453.5,103.694)(6931.5,113.29276150627615)(7409.5,123.2455)(7887.5,133.35723430962344)(8365.5,143.826)(8843.5,154.50561924686193)(9321.5,165.50718828451883)(9799.5,176.844)(10277.5,188.50183891213388)(10755.5,200.51678451882844)(11233.5,212.91894769874477)(11711.5,225.72532635983262)(12189.5,238.95126150627615)(12667.5,252.61902928870293)(13145.5,266.74416527196655)(13623.5,281.3792719665272)(14101.5,296.5673284518829)(14579.5,312.32517364016735)(15057.5,328.69830753138075)(15535.5,345.75547071129705)(16013.5,363.5053430962343)(16491.5,382.03205439330543)(16969.5,401.41702092050207)(17447.5,421.740859832636)(17925.5,443.13734518828454)(18403.5,465.74711715481175)(18881.5,489.6866317991632)(19359.5,515.2018326359832)(19837.5,542.4019079497908)(20315.5,571.3341841004185)(20793.5,602.0434246861925)(21271.5,634.8044414225941)(21749.5,669.9890732217573)(22227.5,707.3027949790795)(22705.5,746.5745418410041)(23183.5,787.7260481171548)(23661.5,830.9119288702929)(24139.5,876.1701464435146)(24617.5,923.622338912134)(25095.5,973.415449790795)(25573.5,1025.5190062761505)(26051.5,1080.0794079497907)(26529.5,1137.2441317991631)(27007.5,1197.1491380753137)(27485.5,1259.997587866109)(27963.5,1325.8901129707112)(28441.5,1395.2120104602511)(28919.5,1468.3164748953975)(29397.5,1545.6936443514644)(29875.5,1627.697179916318)(30353.5,1714.9648619246861)(30831.5,1808.0020460251046)(31309.5,1907.35589958159)(31787.5,2013.702589958159)(32265.5,2127.0940774058577)(32743.5,2248.581280334728)(33221.5,2378.0494142259417)(33699.5,2516.0073828451887)(34177.5,2664.039150627615)(34655.5,2823.221081589958)(35133.5,2993.5568828451883)(35611.5,3175.938089958159)(36089.5,3371.323261506276)(36567.5,3580.968144351464)(37045.5,3805.3915585774057)(37523.5,4045.9028263598325)(38001.5,4303.908370292887)(38479.5,4580.979633891214)(38957.5,4878.916422594142)(39435.5,5201.072190376569)(39913.5,5551.563246861924)(40391.5,5935.1872949790795)(40869.5,6355.800322175732)(41347.5,6819.264579497908)(41825.5,7332.79989330544)(42303.5,7913.116949790795)(42781.5,8584.49539748954)(43259.5,9376.001481171548)(43737.5,10323.55618200837)(44215.5,11494.943769874477)(44693.5,13013.801780334728)(45171.5,15102.628635983265)(45649.5,18314.33333472803)(46127.5,23806.17459414226)(46582.5,34369.86894675926)(46800,43539.844)};
\addlegendentry{Z3seq}

\addplot[name path=pathZ3str3RE , colourZ3str3RE, line width=1.5pt] coordinates {(1,0.003)(242.5,1.0744855371900826)(726.5,3.4655)(1210.5,5.8855)(1694.5,8.3055)(2178.5,10.889551652892562)(2662.5,13.786)(3146.5,16.69)(3630.5,19.594)(4114.5,22.498)(4598.5,25.402)(5082.5,28.306)(5566.5,31.21)(6050.5,34.114)(6534.5,37.018)(7018.5,39.922)(7502.5,42.826)(7986.5,45.73)(8470.5,48.634)(8954.5,51.538)(9438.5,54.442)(9922.5,57.45211570247933)(10406.5,60.8125)(10890.5,64.2005)(11374.5,67.5885)(11858.5,70.9765)(12342.5,74.3645)(12826.5,77.7525)(13310.5,81.1405)(13794.5,84.5285)(14278.5,88.04413636363637)(14762.5,91.898)(15246.5,95.77)(15730.5,99.642)(16214.5,103.60915702479339)(16698.5,107.9315)(17182.5,112.2875)(17666.5,116.68544421487603)(18150.5,121.443)(18634.5,126.283)(19118.5,131.25429338842974)(19602.5,136.5615)(20086.5,141.96053099173554)(20570.5,147.721)(21054.5,153.53384710743802)(21538.5,159.6475)(22022.5,165.9395)(22506.5,172.2315)(22990.5,178.64468388429754)(23474.5,185.4)(23958.5,192.176)(24442.5,199.00360330578513)(24926.5,206.1935)(25410.5,213.4535)(25894.5,220.8263409090909)(26378.5,228.546)(26862.5,236.30813636363635)(27346.5,244.4085)(27830.5,252.6365)(28314.5,261.0807252066116)(28798.5,269.8022272727273)(29282.5,278.8455)(29766.5,288.1148698347107)(30250.5,297.7528615702479)(30734.5,307.75080165289256)(31218.5,318.19113636363636)(31702.5,329.0933388429752)(32186.5,340.5123243801653)(32670.5,352.5468078512397)(33154.5,365.25928305785123)(33638.5,378.71496694214875)(34122.5,392.9674814049587)(34606.5,408.1881095041322)(35090.5,424.496923553719)(35574.5,442.0963760330579)(36058.5,461.27055165289255)(36542.5,482.10873966942154)(37026.5,504.83204132231407)(37510.5,529.7162376033058)(37994.5,557.2993388429752)(38478.5,588.2287933884298)(38962.5,623.190291322314)(39446.5,663.5498512396694)(39930.5,711.7078305785124)(40414.5,771.0276487603306)(40898.5,850.0923057851239)(41382.5,963.3128016528926)(41866.5,1101.6400867768596)(42350.5,1247.2309276859505)(42834.5,1402.3912768595042)(43318.5,1571.1703719008265)(43802.5,1754.472152892562)(44286.5,1953.5502128099172)(44770.5,2173.350842975207)(45254.5,2424.403890495868)(45738.5,2725.6362169421486)(46222.5,3127.2070268595044)(46706.5,3851.115107438017)(47176.5,6425.0990570175445)(47406,11466.323)};
\addlegendentry{Z3str3RE}

\addplot[name path=pathZ3-Trau , colourZ3-Trau, line width=1.5pt] coordinates {(1,0.013)(121.5,1.7120082644628098)(363.5,5.2225)(605.5,8.8525)(847.5,12.548330578512397)(1089.5,16.480061983471074)(1331.5,20.66350826446281)(1573.5,25.187586776859504)(1815.5,30.315636363636365)(2057.5,37.034028925619836)(2299.5,47.26819421487603)(2541.5,63.143694214876035)(2783.5,87.08737190082644)(3025.5,119.03057438016529)(3267.5,152.87779338842975)(3509.5,187.68713223140495)(3751.5,223.38856198347108)(3993.5,259.89136776859505)(4235.5,297.26910743801653)(4477.5,335.72614049586775)(4719.5,375.26788429752065)(4961.5,415.86328512396693)(5203.5,457.628826446281)(5445.5,500.5764338842975)(5687.5,545.049520661157)(5929.5,591.5783884297521)(6171.5,640.2639297520661)(6413.5,691.3767644628099)(6655.5,745.3044504132231)(6897.5,802.9776611570248)(7139.5,864.1297355371901)(7381.5,927.3781404958678)(7623.5,992.0037809917355)(7865.5,1057.7867685950412)(8107.5,1124.6315123966942)(8349.5,1192.5747231404957)(8591.5,1261.5780537190083)(8833.5,1331.6086074380164)(9075.5,1402.7028347107437)(9317.5,1474.9072066115702)(9559.5,1548.5075041322314)(9801.5,1623.6399090909092)(10043.5,1700.4721859504132)(10285.5,1779.2844545454545)(10527.5,1860.0937727272726)(10769.5,1942.9279256198347)(11011.5,2027.9387975206612)(11253.5,2115.4554380165287)(11495.5,2205.679305785124)(11737.5,2299.044983471075)(11979.5,2395.7380041322313)(12221.5,2495.453338842975)(12463.5,2597.9617561983473)(12705.5,2703.040876033058)(12947.5,2811.046743801653)(13189.5,2922.233561983471)(13431.5,3036.8190743801656)(13673.5,3155.0550041322313)(13915.5,3277.2954958677688)(14157.5,3404.018123966942)(14399.5,3535.343376033058)(14641.5,3671.3976983471075)(14883.5,3811.849888429752)(15125.5,3957.2713512396695)(15367.5,4108.5858181818185)(15609.5,4266.528115702479)(15851.5,4431.535669421488)(16093.5,4604.015574380165)(16335.5,4785.366219008264)(16577.5,4975.805739669422)(16819.5,5176.19655785124)(17061.5,5387.800747933884)(17303.5,5612.16417768595)(17545.5,5849.1159669421495)(17787.5,6099.078747933884)(18029.5,6365.81873553719)(18271.5,6652.1750991735535)(18513.5,6961.59579338843)(18755.5,7294.319004132231)(18997.5,7654.666690082645)(19239.5,8048.816743801653)(19481.5,8482.50705785124)(19723.5,8960.417227272726)(19965.5,9492.69403305785)(20207.5,10088.572396694215)(20449.5,10753.385491735537)(20691.5,11507.436082644628)(20933.5,12350.568033057853)(21175.5,13286.226838842975)(21417.5,14340.296863636364)(21659.5,15565.351520661157)(21901.5,17022.022425619834)(22143.5,18795.41270247934)(22385.5,20994.64411570248)(22627.5,23767.618681818185)(22869.5,27349.00535123967)(23111.5,32061.896181818185)(23353.5,38596.88123966942)(23595.5,48042.54463636364)(23737,55329.597317073174)(23759,56566.098)};
\addlegendentry{Z3-Trau}

\path[name path=axisCVC5] (axis cs:0,0) -- (axis cs:44001,0);
\addplot [thick,color=colourCVC5,fill=colourCVC5,fill opacity=0.1] fill between [of=pathCVC5 and axisCVC5];
\path[name path=axisOstrichCEA] (axis cs:0,0) -- (axis cs:49127,0);
\addplot [thick,color=colourOstrichCEA,fill=colourOstrichCEA,fill opacity=0.1] fill between [of=pathOstrichCEA and axisOstrichCEA];
\path[name path=axisOstrich] (axis cs:0,0) -- (axis cs:47370,0);
\addplot [thick,color=colourOstrich,fill=colourOstrich,fill opacity=0.1] fill between [of=pathOstrich and axisOstrich];
\path[name path=axisZ3str3] (axis cs:0,0) -- (axis cs:36317,0);
\addplot [thick,color=colourZ3str3,fill=colourZ3str3,fill opacity=0.1] fill between [of=pathZ3str3 and axisZ3str3];
\path[name path=axisZ3seq] (axis cs:0,0) -- (axis cs:46800,0);
\addplot [thick,color=colourZ3seq,fill=colourZ3seq,fill opacity=0.1] fill between [of=pathZ3seq and axisZ3seq];
\path[name path=axisZ3str3RE] (axis cs:0,0) -- (axis cs:47406,0);
\addplot [thick,color=colourZ3str3RE,fill=colourZ3str3RE,fill opacity=0.1] fill between [of=pathZ3str3RE and axisZ3str3RE];
\path[name path=axisZ3-Trau] (axis cs:0,0) -- (axis cs:23759,0);
\addplot [thick,color=colourZ3-Trau,fill=colourZ3-Trau,fill opacity=0.1] fill between [of=pathZ3-Trau and axisZ3-Trau];
\end{axis}\end{tikzpicture}}

\end{document}