### 2023-04-11 2023-1-19/figures/cactus_plot_heuristic.tex

\documentclass[11pt]{standalone}
\usepackage{color}
\usepackage{tikz}
\usepackage{pgfplots}
\usepgfplotslibrary{fillbetween}
\pgfplotsset{compat=1.16}
\begin{document}
\definecolor{colourOstrichCEA}{HTML}{25333D}
\definecolor{colourUnder Approximation off}{HTML}{0065AB}
\definecolor{colourSimplification off}{HTML}{8939AD}
\definecolor{colourAll off}{HTML}{007E7A}
\resizebox{.95\textwidth}{!}{\pgfplotsset{scaled x ticks=false}\pgfplotsset{scaled y ticks=false}\begin{tikzpicture}\begin{axis}[title=Total,xlabel=Solved instances,ylabel=Time (s),,legend columns=2,legend style={nodes={scale=0.5, transform shape}, fill=none,anchor=east,align=center },axis line style={draw=none}, xtick pos=left, ytick pos=left, ymajorgrids=true, legend style={draw=none},x post scale=2,y post scale=1]
\addplot[name path=pathOstrichCEA , colourOstrichCEA, line width=1.5pt] coordinates {(1,1.256)(251,375.73549500998007)(752,1183.8739021956087)(1253,2022.618491017964)(1754,2877.9200359281435)(2255,3746.0393033932137)(2756,4625.084281437126)(3257,5513.779463073853)(3758,6411.872730538922)(4259,7319.427289421158)(4760,8236.114157684631)(5261,9161.532343313374)(5762,10096.144241516966)(6263,11040.38131736527)(6764,11994.772165668663)(7265,12960.379053892215)(7766,13938.16142714571)(8267,14928.978754491018)(8768,15934.640826347304)(9269,16955.70472255489)(9770,17992.700299401196)(10271,19044.441936127743)(10772,20110.456696606787)(11273,21189.18998203593)(11774,22279.62379241517)(12275,23380.84084031936)(12776,24491.869590818365)(13277,25612.71845309381)(13778,26743.928768463076)(14279,27885.323826347307)(14780,29035.924956087823)(15281,30195.091335329344)(15782,31363.01940518962)(16283,32539.236013972055)(16784,33723.31988023953)(17285,34915.447836327345)(17786,36114.84655888224)(18287,37321.58146307385)(18788,38535.701952095806)(19289,39756.911600798405)(19790,40985.1631017964)(20291,42220.45616367266)(20792,43462.636586826346)(21293,44711.67520558883)(21794,45967.50069461078)(22295,47230.13759281437)(22796,48499.66564471058)(23297,49776.35208782435)(23798,51060.01192814371)(24299,52350.822175648704)(24800,53648.5965748503)(25301,54953.18310379241)(25802,56264.57108582834)(26303,57583.00460678643)(26804,58908.59252295409)(27305,60241.239854291416)(27806,61581.528387225546)(28307,62929.616544910175)(28808,64285.927918163674)(29309,65650.71985229541)(29810,67024.12507584831)(30311,68406.55969061877)(30812,69797.76003592815)(31313,71197.73274650698)(31814,72607.24381437126)(32315,74026.60229740519)(32816,75455.86899201597)(33317,76895.86807984032)(33818,78346.82771656686)(34319,79809.24070259482)(34820,81283.33454291416)(35321,82769.59380638723)(35822,84269.10105988024)(36323,85782.33381836327)(36824,87310.26063073853)(37325,88853.80257085829)(37826,90415.16936327345)(38327,91994.438750499)(38828,93592.4894510978)(39329,95211.69484830339)(39830,96853.58521157685)(40331,98519.82530938124)(40832,100212.75770459081)(41333,101935.10099401198)(41834,103689.06698403193)(42335,105476.94833532933)(42836,107301.12517165668)(43337,109166.71543512975)(43838,111078.39904191616)(44339,113041.1118263473)(44840,115058.31930339322)(45341,117139.44398802395)(45842,119295.3597245509)(46343,121540.35781237525)(46844,123904.0013233533)(47345,126431.7289760479)(47846,129212.59415968064)(48347,132495.7146007984)(48845.5,137664.56463508063)(49095,143594.349)};
\addlegendentry{OstrichCEA}

\addplot[name path=pathUnder Approximation off , colourUnder Approximation off, line width=1.5pt] coordinates {(1,1.143)(247.5,345.5671659919028)(741.5,1093.1268178137652)(1235.5,1881.7518846153846)(1729.5,2688.709206477733)(2223.5,3507.6373157894736)(2717.5,4336.697226720647)(3211.5,5174.640070850202)(3705.5,6020.75543117409)(4199.5,6875.348497975709)(4693.5,7738.697793522268)(5187.5,8610.4509048583)(5681.5,9491.439477732794)(6175.5,10383.11131174089)(6669.5,11286.058060728745)(7163.5,12201.323759109313)(7657.5,13130.214445344129)(8151.5,14072.49958097166)(8645.5,15027.589060728746)(9139.5,15993.548682186234)(9633.5,16968.388783400813)(10127.5,17950.561637651823)(10621.5,18939.66844736842)(11115.5,19935.288463562752)(11609.5,20936.991455465588)(12103.5,21944.461374493927)(12597.5,22957.267400809716)(13091.5,23975.060060728745)(13585.5,24997.827248987855)(14079.5,26025.643408906883)(14573.5,27058.369396761133)(15067.5,28095.932414979758)(15561.5,29138.35586032389)(16055.5,30185.54812145749)(16549.5,31237.71652631579)(17043.5,32294.568991902837)(17537.5,33355.85782591093)(18031.5,34421.65787044534)(18525.5,35492.13017813765)(19019.5,36567.51576923077)(19513.5,37647.78307692308)(20007.5,38732.940694331985)(20501.5,39823.37698582996)(20995.5,40919.24681983805)(21489.5,42020.08232591093)(21983.5,43125.835275303645)(22477.5,44236.932014170045)(22971.5,45353.34651214575)(23465.5,46475.20208299595)(23959.5,47602.54591700405)(24453.5,48735.58559311741)(24947.5,49873.97048987854)(25441.5,51018.06316396761)(25935.5,52167.99957287449)(26429.5,53323.73327327935)(26923.5,54485.43606882591)(27417.5,55653.25192510121)(27911.5,56827.234957489876)(28405.5,58007.57878340081)(28899.5,59194.84101619433)(29393.5,60388.59112955466)(29887.5,61589.210236842104)(30381.5,62796.74455060728)(30875.5,64011.43599797571)(31369.5,65233.93253643725)(31863.5,66464.4293805668)(32357.5,67703.53380364372)(32851.5,68951.35434008097)(33345.5,70208.14511943319)(33839.5,71474.93029149798)(34333.5,72752.14231781376)(34827.5,74040.93050404859)(35321.5,75341.12895748988)(35815.5,76653.30199190283)(36309.5,77978.19206680162)(36803.5,79316.7351194332)(37297.5,80669.18660526315)(37791.5,82036.25991902834)(38285.5,83419.48054655871)(38779.5,84819.95787651822)(39273.5,86240.30562955466)(39767.5,87682.62130364371)(40261.5,89147.46975506074)(40755.5,90638.08723481782)(41249.5,92156.34871457491)(41743.5,93704.68462348178)(42237.5,95286.82339878543)(42731.5,96907.78887246964)(43225.5,98569.67491700404)(43719.5,100277.59814979757)(44213.5,102037.28515789473)(44707.5,103855.00732388663)(45201.5,105743.83414574899)(45695.5,107719.03925101215)(46189.5,109802.37090890687)(46683.5,112039.03551214574)(47177.5,114517.45951214574)(47671.5,117550.44509311742)(48165.5,123028.84814979757)(48419.5,129605.71235714285)(48428,130033.418)};
\addlegendentry{Under Approximation off}

\addplot[name path=pathSimplification off , colourSimplification off, line width=1.5pt] coordinates {(1,1.135)(246.5,363.1045792682927)(738.5,1150.2894695121952)(1230.5,1976.5712662601627)(1722.5,2827.88006097561)(2214.5,3700.4905081300813)(2706.5,4592.667786585365)(3198.5,5505.309475609756)(3690.5,6435.865410569106)(4182.5,7382.667552845528)(4674.5,8345.835542682928)(5166.5,9325.28124186992)(5658.5,10320.03525609756)(6150.5,11330.31531300813)(6642.5,12355.397339430894)(7134.5,13394.562701219513)(7626.5,14448.094024390242)(8118.5,15515.715595528456)(8610.5,16597.635471544716)(9102.5,17694.34049593496)(9594.5,18806.49131504065)(10086.5,19933.432048780487)(10578.5,21074.394237804878)(11070.5,22229.156333333332)(11562.5,23398.27930284553)(12054.5,24582.21882926829)(12546.5,25781.288851626017)(13038.5,26995.324906504065)(13530.5,28224.14231300813)(14022.5,29466.946489837395)(14514.5,30723.627083333333)(15006.5,31995.121766260163)(15498.5,33280.916351626016)(15990.5,34581.29370528455)(16482.5,35896.226640243905)(16974.5,37225.31055691057)(17466.5,38568.79462398374)(17958.5,39926.34044715447)(18450.5,41296.85248577235)(18942.5,42680.20717479674)(19434.5,44077.110611788616)(19926.5,45488.34107113821)(20418.5,46913.13556707317)(20910.5,48352.125798780486)(21402.5,49805.39538211382)(21894.5,51272.918483739835)(22386.5,52754.42153455284)(22878.5,54249.736412601625)(23370.5,55759.46517073171)(23862.5,57283.576719512195)(24354.5,58822.75940243903)(24846.5,60376.9937398374)(25338.5,61946.492164634146)(25830.5,63530.548172764225)(26322.5,65129.932030487806)(26814.5,66745.12398577236)(27306.5,68376.3347296748)(27798.5,70024.90577439025)(28290.5,71690.3053699187)(28782.5,73371.95175406504)(29274.5,75070.203699187)(29766.5,76786.4307703252)(30258.5,78521.07297357723)(30750.5,80274.46002642276)(31242.5,82047.87613211382)(31734.5,83841.58971341464)(32226.5,85655.53554268293)(32718.5,87490.54740447155)(33210.5,89347.99480691057)(33702.5,91229.68758130081)(34194.5,93135.32528455285)(34686.5,95066.21725406504)(35178.5,97024.63288211382)(35670.5,99011.55916463415)(36162.5,101029.76689634146)(36654.5,103081.27141260162)(37146.5,105168.26180081301)(37638.5,107291.90132520326)(38130.5,109453.49715853659)(38622.5,111655.81977845529)(39114.5,113904.8909186992)(39606.5,116202.81763211382)(40098.5,118552.27406707317)(40590.5,120957.2551402439)(41082.5,123425.42828861788)(41574.5,125962.37361585365)(42066.5,128575.37174593496)(42558.5,131275.79281300813)(43050.5,134077.2708373984)(43542.5,136992.64292682928)(44034.5,140037.24509959348)(44526.5,143234.2047682927)(45018.5,146613.52907520323)(45510.5,150227.2450487805)(46002.5,154142.48767479675)(46494.5,158476.08704268292)(46986.5,163481.78608130082)(47478.5,169657.74288617886)(47970.5,179379.5124085366)(48232.5,188736.69675)(48250,189680.595)};
\addlegendentry{Simplification off}

\addplot[name path=pathAll off , colourAll off, line width=1.5pt] coordinates {(1,0.875)(243,239.0925670103093)(728,789.0446762886598)(1213,1427.6660762886597)(1698,2129.867397938144)(2183,2881.0539896907217)(2668,3671.581068041237)(3153,4493.687478350515)(3638,5339.833039175258)(4123,6207.217783505154)(4608,7093.622016494845)(5093,7998.2535546391755)(5578,8919.579428865978)(6063,9857.085991752578)(6548,10809.771478350514)(7033,11776.57105154639)(7518,12756.855435051546)(8003,13750.731303092783)(8488,14758.17447628866)(8973,15778.258529896908)(9458,16810.315026804124)(9943,17854.688107216494)(10428,18910.425841237113)(10913,19977.368713402062)(11398,21055.836604123713)(11883,22146.078154639177)(12368,23247.494826804123)(12853,24359.831055670104)(13338,25483.356445360823)(13823,26618.428564948455)(14308,27765.086074226805)(14793,28923.04323298969)(15278,30092.317987628867)(15763,31272.194832989688)(16248,32463.187107216498)(16733,33665.03872577319)(17218,34878.637705154644)(17703,36103.445179381444)(18188,37338.722492783505)(18673,38585.054059793816)(19158,39842.79948659794)(19643,41112.23856082474)(20128,42392.9905072165)(20613,43685.152076288665)(21098,44988.906665979375)(21583,46304.85870309278)(22068,47632.615063917525)(22553,48972.09419381443)(23038,50323.242482474234)(23523,51686.194051546394)(24008,53061.78827835052)(24493,54449.89010927835)(24978,55849.986849484536)(25463,57262.85256494845)(25948,58688.74006185567)(26433,60128.35171546391)(26918,61580.78451546392)(27403,63046.65391546392)(27888,64526.46483505154)(28373,66020.0550185567)(28858,67528.27430103094)(29343,69051.90326185568)(29828,70591.74819381443)(30313,72149.14522061856)(30798,73724.00676701032)(31283,75316.27581237114)(31768,76926.78957319587)(32253,78555.2702701031)(32738,80202.53150515463)(33223,81871.12324948455)(33708,83561.23319381442)(34193,85273.75785979381)(34678,87009.43252783506)(35163,88769.27956701031)(35648,90556.07208453608)(36133,92373.16419587629)(36618,94221.60716907217)(37103,96102.0720371134)(37588,98018.9705113402)(38073,99975.30944329896)(38558,101974.99814432989)(39043,104021.63824536082)(39528,106116.20065360825)(40013,108263.02482474227)(40498,110468.98961237114)(40983,112746.94890721649)(41468,115107.6707484536)(41953,117562.98611546391)(42438,120126.02667628866)(42923,122802.09602061856)(43408,125607.74125979382)(43893,128581.27053608248)(44378,131755.0405072165)(44863,135193.23917113402)(45348,138956.2646639175)(45833,143194.02182268043)(46318,148200.39157938142)(46803,154703.10293814432)(47274.5,165290.41879475984)(47505,174530.666)};
\addlegendentry{All off}

\path[name path=axisOstrichCEA] (axis cs:0,0) -- (axis cs:49095,0);
\addplot [thick,color=colourOstrichCEA,fill=colourOstrichCEA,fill opacity=0.1] fill between [of=pathOstrichCEA and axisOstrichCEA];
\path[name path=axisUnder Approximation off] (axis cs:0,0) -- (axis cs:48428,0);
\addplot [thick,color=colourUnder Approximation off,fill=colourUnder Approximation off,fill opacity=0.1] fill between [of=pathUnder Approximation off and axisUnder Approximation off];
\path[name path=axisSimplification off] (axis cs:0,0) -- (axis cs:48250,0);
\addplot [thick,color=colourSimplification off,fill=colourSimplification off,fill opacity=0.1] fill between [of=pathSimplification off and axisSimplification off];
\path[name path=axisAll off] (axis cs:0,0) -- (axis cs:47505,0);
\addplot [thick,color=colourAll off,fill=colourAll off,fill opacity=0.1] fill between [of=pathAll off and axisAll off];
\end{axis}\end{tikzpicture}}

\end{document}