### 2023-04-25 2023-1-19/figures/overview_example.tex

  \begin{tikzpicture}[
    shorten >=1pt,node distance=3cm,on grid,>={Stealth[round]},
    initial text=, every state/.style={minimum size = 0.001cm},
    accepting text=$\scriptstyle5\leq r_1\leq 60\wedge 5\leq r_2\leq 60\wedge 0\leq r_3\leq 60\wedge 120 < r_4$, accepting/.style=accepting by arrow,
    accepting where=above
    ]

    \node[state,initial]            (q_0)                      {$q_0$};
    \node[state]                    (q_1) [right=of q_0]       {$q_1$};
    \node[state, accepting]                    (q_2) [right=of q_1]       {$q_2$};

    \path[->] (q_0) edge node [above] {$\scriptstyle \Sigma \setminus a:(1,0,0,1)$} (q_1)
              (q_0) edge [loop below] node  {$\scriptstyle\Sigma \setminus a:(1,0,0,1)$} (q_0)
              (q_1) edge node [above] {$\scriptstyle\Sigma \setminus b:(0,1,0,1)$} (q_2)
              (q_1) edge [loop below] node  {$\scriptstyle\Sigma \setminus b:(0,1,0,1)$} (q_1)
              (q_2) edge [loop below] node  {$\scriptstyle\Sigma \setminus c:(0,0,1,1)$} (q_2);
  \end{tikzpicture}