### 2023-05-04 2023-1-19/figures/cactus_plot_redos.tex

\documentclass[11pt]{standalone}
\usepackage{color}
\usepackage{tikz}
\usepackage{pgfplots}
\usepgfplotslibrary{fillbetween}
\pgfplotsset{compat=1.16}
\begin{document}
\definecolor{colourCVC5}{HTML}{25333D}
\definecolor{colourOstrichCEA}{HTML}{0065AB}
\definecolor{colourOstrich}{HTML}{8939AD}
\definecolor{colourZ3str3}{HTML}{007E7A}
\definecolor{colourZ3seq}{HTML}{CD3517}
\definecolor{colourZ3str3RE}{HTML}{318700}
\definecolor{colourZ3-Trau}{HTML}{80746D}
\resizebox{.95\textwidth}{!}{\pgfplotsset{scaled x ticks=false}\pgfplotsset{scaled y ticks=false}\begin{tikzpicture}\begin{axis}[title=redos,xlabel=Solved instances,ylabel=Time (s),,legend columns=2,legend style={nodes={scale=0.5, transform shape}, fill=none,anchor=east,align=center },axis line style={draw=none}, xtick pos=left, ytick pos=left, ymajorgrids=true, legend style={draw=none},x post scale=2,y post scale=1]
\addplot[name path=pathOstrichCEA , colourOstrichCEA, line width=1.5pt, solid] coordinates {(1,1.268)(8.5,12.6915625)(24.5,34.954875)(40.5,58.2241875)(56.5,82.7331875)(72.5,108.7551875)(88.5,135.7378125)(104.5,163.2775625)(120.5,191.216375)(136.5,219.530625)(152.5,248.17575)(168.5,277.0789375)(184.5,306.2231875)(200.5,335.6075)(216.5,365.1799375)(232.5,394.96425)(248.5,424.948625)(264.5,455.1505)(280.5,485.57425)(296.5,516.2539375)(312.5,547.1195625)(328.5,578.1605)(344.5,609.3965625)(360.5,640.86375)(376.5,672.526125)(392.5,704.3930625)(408.5,736.42525)(424.5,768.6708125)(440.5,801.1625625)(456.5,833.9086875)(472.5,866.889875)(488.5,900.0090625)(504.5,933.2758125)(520.5,966.7259375)(536.5,1000.365375)(552.5,1034.2524375)(568.5,1068.3616875)(584.5,1102.7155625)(600.5,1137.2868125)(616.5,1172.10175)(632.5,1207.1765)(648.5,1242.504)(664.5,1278.005125)(680.5,1313.6974375)(696.5,1349.6119375)(712.5,1385.7789375)(728.5,1422.28125)(744.5,1459.09975)(760.5,1496.2685)(776.5,1533.80525)(792.5,1571.5899375)(808.5,1609.6835)(824.5,1648.163)(840.5,1686.9795625)(856.5,1726.3140625)(872.5,1766.0058125)(888.5,1805.9916875)(904.5,1846.3615625)(920.5,1887.2036875)(936.5,1928.6034375)(952.5,1970.5518125)(968.5,2013.213875)(984.5,2056.5095625)(1000.5,2100.3145625)(1016.5,2144.6456875)(1032.5,2189.60025)(1048.5,2235.2554375)(1064.5,2281.8241875)(1080.5,2329.6734375)(1096.5,2379.1441875)(1112.5,2429.9625)(1128.5,2481.816375)(1144.5,2534.979875)(1160.5,2589.3324375)(1176.5,2645.0780625)(1192.5,2701.892125)(1208.5,2759.7080625)(1224.5,2818.4853125)(1240.5,2878.14725)(1256.5,2938.6613125)(1272.5,2999.7675)(1288.5,3061.7365)(1304.5,3124.550125)(1320.5,3188.13575)(1336.5,3252.6558125)(1352.5,3318.13175)(1368.5,3384.681875)(1384.5,3451.851)(1400.5,3519.970625)(1416.5,3589.426375)(1432.5,3660.527)(1448.5,3733.2529375)(1464.5,3807.681625)(1480.5,3883.4769375)(1496.5,3961.36525)(1512.5,4041.9930625)(1528.5,4126.6176875)(1544.5,4219.805375)(1560.5,4326.948125)(1576.5,4449.7968125)(1592.5,4610.8235625)(1601.5,4744.1385)(1604,4795.597)};
\addlegendentry{OstrichCEA}

\addplot[name path=pathOstrich , colourOstrich, line width=1.5pt, dashdotted] coordinates {(1,1.042)(7.5,10.421857142857142)(21.5,29.635571428571428)(35.5,50.016357142857146)(49.5,71.21285714285715)(63.5,93.25821428571429)(77.5,115.84392857142856)(91.5,138.86014285714288)(105.5,162.4882142857143)(119.5,186.6072142857143)(133.5,211.14664285714287)(147.5,236.0525714285714)(161.5,261.4626428571429)(175.5,287.39128571428574)(189.5,313.71728571428577)(203.5,340.4631428571428)(217.5,367.5703571428572)(231.5,395.1613571428572)(245.5,423.17785714285714)(259.5,451.473)(273.5,480.01521428571425)(287.5,508.8493571428572)(301.5,537.9797142857143)(315.5,567.4725)(329.5,597.379)(343.5,627.7412857142857)(357.5,658.6113571428572)(371.5,689.8423571428572)(385.5,721.4651428571428)(399.5,753.5415)(413.5,786.0342857142857)(427.5,818.8502857142856)(441.5,852.0047857142857)(455.5,885.4851428571428)(469.5,919.4435)(483.5,953.7673571428571)(497.5,988.4813571428572)(511.5,1023.6364285714285)(525.5,1059.1977857142856)(539.5,1095.152357142857)(553.5,1131.447857142857)(567.5,1168.1000714285713)(581.5,1205.2960714285714)(595.5,1243.024642857143)(609.5,1281.2957857142858)(623.5,1320.0607857142857)(637.5,1359.327357142857)(651.5,1398.982)(665.5,1439.111642857143)(679.5,1479.8170714285714)(693.5,1521.0342857142857)(707.5,1562.7634285714287)(721.5,1605.0646428571429)(735.5,1647.8864285714287)(749.5,1691.2777857142858)(763.5,1735.378642857143)(777.5,1779.983642857143)(791.5,1825.111)(805.5,1870.7785714285715)(819.5,1917.3739285714287)(833.5,1964.793142857143)(847.5,2013.078857142857)(861.5,2062.0352857142857)(875.5,2111.9066428571427)(889.5,2162.6124285714286)(903.5,2214.2143571428574)(917.5,2266.915285714286)(931.5,2320.8507142857143)(945.5,2375.982571428571)(959.5,2432.2053571428573)(973.5,2489.9445)(987.5,2549.594642857143)(1001.5,2611.896214285714)(1015.5,2677.7935714285713)(1029.5,2746.8089285714286)(1043.5,2819.2254285714284)(1057.5,2895.4688571428574)(1071.5,2976.7548571428574)(1085.5,3063.5255714285713)(1099.5,3156.2792142857143)(1113.5,3256.008714285714)(1127.5,3361.891214285714)(1141.5,3475.924785714286)(1155.5,3596.808)(1169.5,3725.2726428571427)(1183.5,3863.3013571428573)(1197.5,4012.4511428571427)(1211.5,4173.714785714286)(1225.5,4349.4121428571425)(1239.5,4536.0506428571425)(1253.5,4733.094928571428)(1267.5,4940.848285714285)(1281.5,5164.210285714285)(1295.5,5406.3664285714285)(1309.5,5666.005357142857)(1323.5,5947.754357142857)(1337.5,6272.523357142857)(1351.5,6649.797214285714)(1365.5,7108.828785714285)(1377.5,7661.63)(1384,7972.953)};
\addlegendentry{Ostrich}

\addplot[name path=pathCVC5 , colourCVC5, line width=1.5pt, dashed, dash pattern=on 2pt off 8pt] coordinates {(1,0.007)(7,0.08584615384615384)(20,0.305)(33,0.6530769230769231)(46,1.1311538461538462)(59,1.6853846153846155)(72,2.2931538461538463)(85,2.9566923076923075)(98,3.686230769230769)(111,4.479461538461538)(124,5.35023076923077)(137,6.312153846153846)(150,7.359153846153846)(163,8.470615384615385)(176,9.657692307692308)(189,10.93923076923077)(202,12.307615384615385)(215,13.752153846153846)(228,15.272)(241,16.85884615384615)(254,18.51792307692308)(267,20.254153846153848)(280,22.07076923076923)(293,24.015)(306,26.139692307692307)(319,28.53876923076923)(332,31.162307692307692)(345,34.02146153846154)(358,37.16815384615384)(371,40.49684615384616)(384,44.06376923076923)(397,47.8823076923077)(410,51.92207692307692)(423,56.27353846153846)(436,60.88161538461538)(449,65.64184615384616)(462,70.59938461538461)(475,75.76446153846153)(488,81.06407692307693)(501,86.49453846153847)(514,92.05415384615384)(527,97.75361538461539)(540,103.56776923076923)(553,109.49507692307692)(566,115.5783846153846)(579,121.86215384615384)(592,128.30061538461538)(605,134.89915384615384)(618,141.65592307692307)(631,148.56823076923078)(644,155.6096923076923)(657,162.78415384615386)(670,170.11546153846155)(683,177.60553846153846)(696,185.30261538461536)(709,193.18292307692306)(722,201.2533076923077)(735,209.55253846153846)(748,218.15223076923078)(761,227.18307692307692)(774,236.63753846153847)(787,246.61015384615385)(800,257.15207692307695)(813,268.3446923076923)(826,280.29207692307693)(839,293.0210769230769)(852,306.7554615384616)(865,321.323)(878,336.84784615384615)(891,353.35807692307696)(904,370.93038461538464)(917,389.3006923076923)(930,408.32023076923076)(943,428.23615384615385)(956,449.51346153846157)(969,472.04553846153846)(982,496.191)(995,523.0977692307692)(1008,553.4396923076922)(1021,586.0476923076923)(1034,621.8947692307693)(1047,661.3595384615385)(1060,704.5181538461538)(1073,751.4005384615385)(1086,802.835)(1099,862.5549230769232)(1112,931.1657692307692)(1125,1009.1157692307693)(1138,1101.7934615384615)(1151,1230.164076923077)(1164,1424.009923076923)(1177,1703.8506923076923)(1190,2059.8683846153845)(1203,2464.8026923076923)(1216,2921.014769230769)(1229,3424.7666153846153)(1241.5,3993.243)(1249,4346.266)};
\addlegendentry{CVC5}

\addplot[name path=pathZ3str3 , colourZ3str3, line width=1.5pt, dashed, dash pattern=on 8pt off 2pt] coordinates {(1,0.017)(5,0.13355555555555554)(14,0.37611111111111106)(23,0.639)(32,0.9286666666666666)(41,1.2483333333333333)(50,1.5871111111111111)(59,1.942777777777778)(68,2.3183333333333334)(77,2.710444444444444)(86,3.1183333333333336)(95,3.5474444444444444)(104,3.9951111111111115)(113,4.454666666666667)(122,4.926111111111111)(131,5.410111111111111)(140,5.907444444444445)(149,6.420333333333333)(158,6.947111111111111)(167,7.487666666666667)(176,8.041777777777778)(185,8.613666666666665)(194,9.2)(203,9.802666666666665)(212,10.42)(221,11.05711111111111)(230,11.704777777777778)(239,12.366333333333333)(248,13.042)(257,13.737)(266,14.446666666666665)(275,15.176222222222222)(284,15.924)(293,16.69366666666667)(302,17.484333333333332)(311,18.304)(320,19.147333333333332)(329,20.019777777777776)(338,20.91433333333333)(347,21.834222222222223)(356,22.779333333333334)(365,23.74577777777778)(374,24.738777777777777)(383,25.758666666666667)(392,26.808777777777777)(401,27.893666666666668)(410,29.01788888888889)(419,30.171666666666667)(428,31.351)(437,32.574)(446,33.85466666666667)(455,35.18366666666667)(464,36.55377777777778)(473,37.970888888888894)(482,39.43866666666666)(491,40.962111111111106)(500,42.56055555555555)(509,44.21877777777778)(518,45.95133333333334)(527,47.779111111111106)(536,49.68177777777778)(545,51.67422222222222)(554,53.775333333333336)(563,55.97422222222222)(572,58.306666666666665)(581,60.77277777777778)(590,63.406666666666666)(599,66.23022222222222)(608,69.23144444444443)(617,72.49055555555556)(626,76.05744444444444)(635,79.90144444444444)(644,84.237)(653,88.88711111111111)(662,94.08977777777778)(671,99.92533333333333)(680,106.4301111111111)(689,114.00422222222223)(698,122.53)(707,132.107)(716,143.86844444444444)(725,158.1821111111111)(734,174.91488888888887)(743,194.845)(752,218.70155555555556)(761,248.3372222222222)(770,290.79277777777776)(779,348.5783333333333)(788,417.857)(797,510.13644444444446)(806,641.939)(815,831.2733333333334)(824,1073.363111111111)(833,1364.137888888889)(842,1704.5152222222223)(851,2110.631888888889)(858,2475.6972)(862,2647.177)};
\addlegendentry{Z3str3}

\addplot[name path=pathZ3seq , colourZ3seq, line width=1.5pt, dashed, dash pattern=on 1pt off 2pt] coordinates {(1,0.015)(7,0.13392307692307692)(20,0.4753076923076923)(33,1.0751538461538461)(46,2.187769230769231)(59,3.741076923076923)(72,5.534846153846154)(85,7.5522307692307695)(98,9.893307692307692)(111,12.54646153846154)(124,15.76876923076923)(137,19.567153846153847)(150,23.959615384615383)(163,28.68123076923077)(176,33.70907692307692)(189,39.13515384615384)(202,45.02061538461538)(215,51.28153846153846)(228,57.95976923076923)(241,64.94292307692308)(254,72.15346153846153)(267,79.66346153846153)(280,87.47692307692307)(293,95.62438461538461)(306,104.04515384615384)(319,112.723)(332,121.75030769230769)(345,131.2003076923077)(358,140.9993076923077)(371,151.17884615384617)(384,161.71284615384616)(397,172.62730769230768)(410,184.0403076923077)(423,195.81446153846153)(436,208.03846153846152)(449,220.69846153846154)(462,233.92207692307693)(475,247.7913076923077)(488,262.59430769230767)(501,278.44446153846155)(514,295.05923076923074)(527,312.72138461538464)(540,331.90353846153846)(553,352.35630769230767)(566,373.94807692307694)(579,396.50646153846156)(592,419.9123846153846)(605,444.309)(618,470.0917692307693)(631,497.68453846153847)(644,526.4123076923078)(657,557.0483846153846)(670,590.1912307692307)(683,624.9178461538461)(696,661.1273076923078)(709,699.3653076923077)(722,740.2164615384615)(735,783.0497692307692)(748,827.2873846153847)(761,873.849)(774,923.3921538461539)(787,974.9501538461539)(800,1029.0558461538462)(813,1086.2576153846155)(826,1147.4029230769231)(839,1212.1193846153844)(852,1281.6894615384615)(865,1356.7432307692306)(878,1437.5274615384615)(891,1525.3787692307692)(904,1622.077923076923)(917,1726.5867692307693)(930,1841.5621538461537)(943,1966.1236923076922)(956,2097.5923076923077)(969,2237.991846153846)(982,2388.815846153846)(995,2550.726153846154)(1008,2724.7893076923074)(1021,2912.461)(1034,3108.7013846153845)(1047,3318.6526923076926)(1060,3540.098)(1073,3780.940923076923)(1086,4046.5943846153846)(1099,4341.672692307692)(1112,4661.215)(1125,5005.762769230769)(1138,5382.683384615385)(1151,5793.749307692308)(1164,6244.779692307692)(1177,6752.641307692308)(1190,7313.696153846154)(1203,7921.302461538462)(1216,8582.782846153847)(1226,9128.86)(1231,9358.232)};
\addlegendentry{Z3seq}

\addplot[name path=pathZ3str3RE , colourZ3str3RE, line width=1.5pt, dashed, dash pattern=on 2pt off 4pt] coordinates {(1,0.007)(8.5,0.0758125)(24.5,0.263125)(40.5,0.515125)(56.5,0.7930625)(72.5,1.0844375)(88.5,1.3903125)(104.5,1.71125)(120.5,2.04725)(136.5,2.397)(152.5,2.7555625)(168.5,3.12575)(184.5,3.508)(200.5,3.9005)(216.5,4.30275)(232.5,4.7171875)(248.5,5.1435)(264.5,5.580375)(280.5,6.032125)(296.5,6.4955)(312.5,6.9651875)(328.5,7.4506875)(344.5,7.9521875)(360.5,8.4674375)(376.5,8.9958125)(392.5,9.537)(408.5,10.0895)(424.5,10.658)(440.5,11.238875)(456.5,11.83225)(472.5,12.4380625)(488.5,13.0555)(504.5,13.6829375)(520.5,14.3269375)(536.5,14.9929375)(552.5,15.680375)(568.5,16.39425)(584.5,17.128375)(600.5,17.8755)(616.5,18.6303125)(632.5,19.3979375)(648.5,20.1785625)(664.5,20.972)(680.5,21.78225)(696.5,22.6196875)(712.5,23.4789375)(728.5,24.355875)(744.5,25.24875)(760.5,26.1598125)(776.5,27.08725)(792.5,28.0455625)(808.5,29.0368125)(824.5,30.0645625)(840.5,31.1379375)(856.5,32.2518125)(872.5,33.4005)(888.5,34.5898125)(904.5,35.821375)(920.5,37.097625)(936.5,38.4284375)(952.5,39.8054375)(968.5,41.2453125)(984.5,42.7475)(1000.5,44.3534375)(1016.5,46.0835)(1032.5,47.924875)(1048.5,49.8915625)(1064.5,52.0588125)(1080.5,54.4555625)(1096.5,57.04125)(1112.5,59.751375)(1128.5,62.618375)(1144.5,65.6106875)(1160.5,68.7401875)(1176.5,72.18275)(1192.5,75.974)(1208.5,80.25575)(1224.5,85.5225625)(1240.5,91.56425)(1256.5,98.3388125)(1272.5,106.0435)(1288.5,114.47225)(1304.5,123.289375)(1320.5,132.7138125)(1336.5,142.9955625)(1352.5,154.5948125)(1368.5,168.214)(1384.5,184.7535625)(1400.5,207.01025)(1416.5,235.7268125)(1432.5,270.52575)(1448.5,318.257)(1464.5,393.0419375)(1480.5,494.505875)(1496.5,608.450625)(1512.5,744.9885)(1528.5,957.6901875)(1544.5,1302.16275)(1560.5,1835.9473125)(1570.5,2309.625)(1574,2451.289)};
\addlegendentry{Z3str3RE}

\addplot[name path=pathZ3-Trau , colourZ3-Trau, line width=1.5pt, dashed, dash pattern=on 4pt off 2pt] coordinates {(1,0.016)(2,0.053)(5,0.111)(8,0.17233333333333334)(11,0.24133333333333334)(14,0.35633333333333334)(17,0.532)(20,0.7983333333333333)(23,1.095)(26,1.5226666666666668)(29,2.1523333333333334)(32,2.9396666666666667)(35,3.9976666666666665)(38,5.294666666666667)(41,6.714)(44,8.258666666666667)(47,9.852333333333334)(50,11.533666666666665)(53,13.339333333333334)(56,15.281333333333334)(59,17.312)(62,19.477666666666668)(65,21.76233333333333)(68,24.18)(71,26.796)(74,29.68733333333333)(77,32.67433333333333)(80,35.758)(83,38.97)(86,42.28933333333334)(89,45.894333333333336)(92,49.85166666666667)(95,53.928333333333335)(98,58.059)(101,62.247)(104,66.66433333333333)(107,71.34166666666667)(110,76.268)(113,81.44766666666668)(116,86.93366666666667)(119,92.68)(122,98.49866666666667)(125,104.398)(128,110.51466666666667)(131,116.94566666666667)(134,123.67666666666668)(137,131.009)(140,138.821)(143,146.862)(146,155.07933333333335)(149,163.47833333333335)(152,172.06766666666667)(155,180.78866666666664)(158,190.297)(161,200.83833333333334)(164,211.90466666666666)(167,223.99133333333333)(170,236.898)(173,250.152)(176,263.98466666666667)(179,278.498)(182,293.6046666666667)(185,309.477)(188,326.655)(191,344.77)(194,363.702)(197,382.9943333333333)(200,402.6673333333333)(203,422.945)(206,444.157)(209,467.0073333333333)(212,491.593)(215,517.5033333333333)(218,544.213)(221,571.7546666666666)(224,601.098)(227,631.7703333333334)(230,663.7593333333334)(233,697.4513333333334)(236,732.3156666666666)(239,768.4563333333334)(242,805.529)(245,843.917)(248,883.4013333333334)(251,926.3146666666667)(254,974.039)(257,1029.1116666666667)(260,1089.092)(263,1154.1753333333334)(266,1228.3816666666667)(269,1306.6253333333332)(272,1392.2936666666667)(275,1481.7986666666668)(278,1575.6423333333332)(281,1676.5096666666668)(284,1782.0873333333332)(287,1895.8256666666668)(290,2017.9586666666667)(293,2152.2976666666664)(296,2303.433)(299,2464.662)(302,2630.6766666666667)(305,2799.423)(308,2973.517)(311,3092.419)};
\addlegendentry{Z3-Trau}

\path[name path=axisOstrichCEA] (axis cs:0,0) -- (axis cs:1604,0);
\addplot [thick,color=colourOstrichCEA,fill=colourOstrichCEA,fill opacity=0.1] fill between [of=pathOstrichCEA and axisOstrichCEA];
\path[name path=axisOstrich] (axis cs:0,0) -- (axis cs:1384,0);
\addplot [thick,color=colourOstrich,fill=colourOstrich,fill opacity=0.1] fill between [of=pathOstrich and axisOstrich];
\path[name path=axisCVC5] (axis cs:0,0) -- (axis cs:1249,0);
\addplot [thick,color=colourCVC5,fill=colourCVC5,fill opacity=0.1] fill between [of=pathCVC5 and axisCVC5];
\path[name path=axisZ3str3] (axis cs:0,0) -- (axis cs:862,0);
\addplot [thick,color=colourZ3str3,fill=colourZ3str3,fill opacity=0.1] fill between [of=pathZ3str3 and axisZ3str3];
\path[name path=axisZ3seq] (axis cs:0,0) -- (axis cs:1231,0);
\addplot [thick,color=colourZ3seq,fill=colourZ3seq,fill opacity=0.1] fill between [of=pathZ3seq and axisZ3seq];
\path[name path=axisZ3str3RE] (axis cs:0,0) -- (axis cs:1574,0);
\addplot [thick,color=colourZ3str3RE,fill=colourZ3str3RE,fill opacity=0.1] fill between [of=pathZ3str3RE and axisZ3str3RE];
\path[name path=axisZ3-Trau] (axis cs:0,0) -- (axis cs:311,0);
\addplot [thick,color=colourZ3-Trau,fill=colourZ3-Trau,fill opacity=0.1] fill between [of=pathZ3-Trau and axisZ3-Trau];
\end{axis}\end{tikzpicture}}

\end{document}