### 2023-05-13 2023-1-19/tables/table_summary.tex
add missing files

%\documentclass[11pt]{standalone}
%\usepackage{color}
%\usepackage{tikz}
%\usepackage{pgfplots}
%\usepgfplotslibrary{fillbetween}
%\pgfplotsset{compat=1.16}
%\begin{document}
\resizebox{.95\textwidth}{!}{\begin{tabular}{|c |c |c |c |c |c |c |c |}
    \hline
         solver                                                                  & CVC5       & Z3seq      & Z3-Trau & Z3str3     & Z3str3RE   & OSTRICH    & OSTRICH$^{\rm RECL}$ \\
    \hline\hline
    solved correctly                                                         & 44,704     & 46,426     & 24,121  & 35,866     & 47,607     & 41,606      & {\bf 48,698}         \\
    \hline
    \begin{tabular}{c}average time \\ per instance \\ (seconds)\end{tabular} & 5.67       & 3.49       & 5.41    & 8.54       & {\bf 1.71} & 3.9        & 2.93                 \\
    \hline
  \end{tabular}}

%\end{document}