### 2023-10-20 2023-1-19/crv/sections/conclusion.tex

%!TEX root = ../main.tex

This work proposed an efficient automata-theoretical approach for solving string constraints with regex-counting and string-length. The approach is based on encoding counting operators in regular expressions by cost registers symbolically instead of unfolding them explicitly. Moreover, this work proposed automata-size reduction techniques to improve performance further. Finally, we used two benchmark suites comprising 48,843 instances in total to evaluate the performance of our approach. The experimental results show that our approach can solve more instances than the state-of-the-art best solvers, at a comparable or faster speed, especially when the counting and length bounds are large. 
For the future work, we plan to investigate how the symbolic approach can be extended to reason about nested counting operators.