### 2024-06-02 2024-5-13/sections/conclusion.tex
fix the syntax error in conclusion

%!TEX root = ../main.tex

This work proposed an efficient automata-theoretical approach for solving string constraints with regex-counting and string-length. The approach is based on encoding counting operators in regular expressions by cost registers symbolically instead of unfolding them explicitly. Moreover, this work proposed various optimization techniques to improve the performance further. Finally, we used three benchmark suites comprising 49,843 instances in total to evaluate the performance of our approach. The experimental results show that our approach can solve more instances than the best state-of-the-art solver, at a comparable or faster speed, especially when the counting and length bounds are large or when counting operators are nested with some other counting operators or complement operators. 
%For the future work, we plan to investigate how the symbolic approach can be extended to reason about nested counting operators.