### 2024-12-31 2024-5-13/sections/algorithms.tex
Some revision from prof. wu and change about complexity analysis

%!TEX root = ../main.tex

The goal of this section is to show how to solve RECL constraints by utilizing CEFA. 
At first, we reduce the satisfiability of RECL constraints to a decision problem defined in the sequel. Then we propose a decision procedure for this problem.

\begin{definition}[$\cefadec$]\label{def-cefadec}
Let $x_1, \cdots, x_n$ be string variables, $\Lambda_{x_1}, \cdots, \Lambda_{x_n}$ be nonempty sets of CEFA over the alphabet $\Sigma$ with $\Lambda_{x_i} = \{\aut_{i, 1}, \cdots, \aut_{i, l_i}\}$ for every $i \in [n]$ where the sets of registers $R_{\aut_{1, 1}}, \cdots, R_{\aut_{1, l_1}}, \cdots, R_{\aut_{n, 1}}, \cdots, R_{\aut_{n, l_n}}$ are mutually disjoint, moreover, let $\varphi$ be an LIA formula whose free variables are from $ \bigcup \limits_{i \in [n], j \in [l_i]} R_{\aut_{i, j}}$. Then  the CEFA in $\Lambda_{x_1}, \cdots, \Lambda_{x_n}$ are said to be nonempty w.r.t. $\varphi$ if there are assignments $\theta: \{x_1, \cdots, x_n\} \rightarrow \Sigma^*$ and vectors $\myvec{v}_{i,j}$
such that $(\theta(x_i); \myvec{v}_{i,j}) \in \Lang(\aut_{i, j})$ and  $\varphi[(\myvec{v}_{i,j}/R_{\aut_{i,j}})]$ is true, for every $i \in [n], j \in [l_i]$.
\end{definition}

\begin{proposition}[\cite{atva2020}]\label{prop-cefadec}
$\cefadec$ is PSPACE-complete. 
\end{proposition}
%\begin{proof}
%  The lower bound follows from the fact that the non-emptiness problem of the intersection of NFAs is PSPACE-complete \cite{pspace_int}. For the upper bound, \cite{atva2020} showed that $\cefadec$ is in PSPACE by reducing it to monotonic counter machines.
%\end{proof}
Note that the decision procedure in \cite{atva2020} was only used to prove the upper bound in Proposition~\ref{prop-cefadec} and not implemented as a matter of fact. Instead, the symbolic model checker nuXmv~\cite{nuxmv} was used to solve $\cefadec$. We do not rely on nuXmv in this work and shall propose a new algorithm for solving $\cefadec$ in Section~\ref{subsec:cefadec}.

\subsection{From $\reclsat$ to $\cefadec$} \label{subsec:regex2cefa}

Let $\varphi$ be a RECL constraint and $x_1, \cdots, x_n$ be an enumeration of the string variables occurring in $\varphi$. Moreover, let $\varphi \equiv \varphi_1 \wedge \varphi_2$ such that $\varphi_1$ is a conjunction of regular membership constraints of $\varphi$, and $\varphi_2$ is a conjunction of length constraints of $\varphi$.
We shall reduce the satisfiability of $\varphi$ to an instance of $\cefadec$.

At first, we show how to construct a CEFA from a regex where counting operators may occur. \hdh{Our approach largely follows Thompson's construction~\cite{Thompson1968} but we avoid using the $\epsilon$-transitions. Similar ideas of representing counting operators as counters in automata have been used in e.g. \cite{redos_lenka}. }
%\hdh{and we remove $\epsilon$-transitions on the fly (See union at Section \ref{sec:automaton} and construction of counting operators below)}. 
%\hdh{Some other algorithms to construct NFA from regular expressions are also available, for instance, Glushkov’s construction~\cite{Gluskov61} and its variants~\cite{redos_lenka,HS+23,GKM23}.}

Let us start with register-representable regexes defined in the sequel.

Let us fix an alphabet $\Sigma$.

Let $e$ be a regex over $\Sigma$. Then an occurrence of counting operators in $e$, say $(e')^{\{m,n\}}$ (or $(e')^{\{m, \infty\}}$), is said to be \emph{register-representable} if $(e')^{\{m,n\}}$ (or $(e')^{\{m, \infty\}}$) is not in the scope of a Kleene star, another counting operator, complement, or language difference in $e$. We say that $e$ is \emph{register-representable} if and only if all occurrences of counting operators in $e$ are register-representable. For instance, $a^{\{2, 6\}} \cap a^{\{4, \infty\}}$ is register-representable, while $\overline{a^{\{2, 6\}}}$ and $(a^{\{2, 6\}})^{\{4,\infty\}}$ are not since $a^{\{2,6\}}$ is in the scope of complement and the counter operator $\{4, \infty\}$ respectively.

Let $e$ be a register-representable regex over $\Sigma$. By the following procedure, we will construct a CEFA out of $e$, denoted by $\aut_e$. 
\begin{enumerate}
%
\item For each sub-expression $(e')^{\{m,n\}}$ with $m \le n$ (resp. $(e')^{\{m, \infty\}}$) of $e$, we construct a CEFA $\aut_{(e')^{\{m,n\}}}$ (resp. $\aut_{(e')^{\{m, \infty\}}}$). Let $\aut_{e'} = (Q, \Sigma, \delta, I, F)$.  
%
Then $\aut_{(e')^{\{m,n\}}} = ((r'), Q', \Sigma, \delta'', I', F', \alpha')$, where $r'$ is a new register, $Q' = Q \cup \{q_0\}$ with $q_0 \not \in Q$, $I' = \{q_0\}$, $F'= F \cup \{q_0\}$, and 
%
$$
\begin{array}{l c l}
\delta'' & = & \{(q, a, q', (0)) \mid (q, a, q') \in \delta\}\  \cup \\
& & \{(q_0, a, q', (1))  \mid \exists q'_0 \in I.\ (q'_0, a, q') \in \delta\}\ \cup \\
& & \{(q, a, q', (1)) \mid q \in F, \exists q'_0 \in I.\ (q'_0, a, q') \in \delta\},
\end{array}
$$ 
%
moreover, $\alpha' = m \le r' \le n$ if $I \cap F = \emptyset$, otherwise $\alpha' = r' \le n$ .  (Intuitively, if $\varepsilon$ is accepted by $\aut_{e'}$, then the value of $r'$ can be less than $m$.) Moreover, $\aut_{(e')^{\{m, \infty\}}}$ is constructed by adapting $\alpha'$ in $\aut_{(e')^{\{m,n\}}}$ as follows: $\alpha' = m \le r'$ if $I \cap F = \emptyset$ and $\alpha' = \ltrue$ otherwise. 
%

\item For each sub-expression $e'$ of $e$ such that $e'$ contains occurrences of counting operators but $e'$ itself is not of the form $(e'_1)^{\{m,n\}}$ or $(e'_1)^{\{m,\infty\}}$, from the assumption that $e$ is register-representable, we know that $e'$ is of the form $e'_1 \concat e'_2$, $e'_1 + e'_2$, $e'_1 \cap e'_2$, or $(e'_1)$. For $e' = (e'_1)$, we have $\aut_{e'} = \aut_{e'_1}$. For $e' = e'_1 \concat e'_2$, $e' = e'_1 + e'_2$, or $e' = e'_1 \cap e'_2$, suppose that CEFA $\aut_{e'_1}$ and $\aut_{e'_2}$ have been constructed.

\item For each maximal sub-expression $e'$ of $e$ such that $e'$ contains no occurrences of counting operators, an NFA $\aut_{e'}$ can be constructed by structural induction on the syntax of $e'$. 
Then we have $\aut_{e'} = \aut_{e'_1} \concat \aut_{e'_2}$, $\aut_{e'} = \aut_{e'_1} \cup \aut_{e'_2}$, or $\aut_{e'} = \aut_{e'_1} \cap \aut_{e'_2}$. 
\end{enumerate}

For non-register-representable regexes, we first convert them into register-representable regexes by unfolding all the non-register-representable occurrences of counting operators. After that, we utilize the aforementioned procedure to construct CEFA. For instance, $(a^{\{2, 6\}} \concat b^*)^{\{2,\infty\}}$ is transformed into $(aa(\varepsilon + a + aa + aaa+aaaa) \concat b^*)^{\{2, \infty\}}$. 
%$(a^{\{2, 6\}} \concat b^*) \setminus (a + b)^{\{4, \infty\}}$ is transformed into $(aa(\varepsilon + a + aa + aaa + aaaa) \concat b^*) \setminus (a+b)(a+b)(a+b)(a+b)(a+b)^*$.

For each $i \in [n]$, let $x_i \in e_{i, 1}, \cdots, x_i \in e_{i, l_i}$ be an enumeration of the regular membership constraints for $x_i$ in $\varphi_1$.  Then we can construct CEFA $\aut_{i, j}$ from $e_{i, j}$ for each $i \in [n]$ and $j \in [l_i]$. Moreover, we construct another CEFA $\aut_{i, 0}$ for each $i \in [n]$ to model the length of $x_i$. Specifically, $\aut_{i,0}$ is constructed as $((r_{i,0}), \{q_{i,0}\}, \Sigma, \delta_{i,0}, \{q_{i,0}\}, \{q_{i,0}\}, \ltrue)$ where $r_{i,0}$ is a fresh register and $\delta_{i,0} = \{(q_{i,0}, a, q_{i,0}, (1)) \mid a \in \Sigma\}$. 
Let $\Lambda_{x_i} = \{\aut_{i,0}, \aut_{i, 1}, \cdots, \aut_{i, l_i}\}$ for each $i \in [n]$, and $\varphi'_2 \equiv \varphi_2[r_{1,0}/|x_1|, \cdots, r_{n,0}/|x_n|]$. Then the satisfiability of $\varphi$ is reduced to the non-emptiness of CEFAs in $\Lambda_{x_1}, \cdots, \Lambda_{x_n}$ w.r.t. $\varphi'_2$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Solving $\cefadec$} \label{subsec:cefadec}

In this section, we present a procedure to solve the $\cefadec$ problem: Suppose that $x_1, \cdots, x_n$ are mutually distinct string variables, $\Lambda_{x_1}, \cdots, \Lambda_{x_n}$ are nonempty sets of CEFA over the same alphabet $\Sigma$ where $\Lambda_{x_i} = \{\aut_{i, 1}, \cdots,$ $ \aut_{i, l_i}\}$ for every $i \in [n]$. Moreover, the sets of registers $R_{\aut_{1, 1}}, \cdots, R_{\aut_{n, l_n}}$ are mutually disjoint, and $\varphi$ is a LIA formula whose free variables are from $\bigcup \limits_{i \in [n], j \in [l_i]} R_{\aut_{i, j}}$. 
The procedure comprises three steps.

\medskip
\noindent
\emph{Step 1 (Computing intersection automata).} For each $i \in [n]$, compute a CEFA $\cB_{i} = \aut_{i,1} \cap \cdots \cap \aut_{i, l_i}$, and let $\Lambda'_{x_i} := \{\cB_i\}$. \qed

After Step 1, the non-emptiness of CEFAs in $\Lambda_{x_1}, \cdots, \Lambda_{x_n}$ w.r.t. $\varphi$ is reduced to the non-emptiness of CEFAs in $\Lambda'_{x_1}, \cdots, \Lambda'_{x_n}$ w.r.t. $\varphi$.

In the following steps, we reduce the non-emptiness of CEFAs in $\Lambda'_{x_1}, \cdots, $ $\Lambda'_{x_n}$ w.r.t. $\varphi$ to the satisfiability of an LIA formula. The reduction relies on the following two observations. 
%The first observation follows from the fact that $\Lambda'_{x_i}$ for each $i \in [n]$ contains exactly one CEFA, and the second observation follows from the fact that the non-emptiness of exactly one CEFA w.r.t an LIA formula is entirely unrelated to the alphabet.

\begin{observation}\label{obs-cefa-output}
CEFA in $\Lambda'_{x_1} = \{\cB_1\}, \cdots, \Lambda'_{x_n} = \{\cB_n\}$ are nonempty w.r.t. $\varphi$ iff there are \hdh{outputs} $\myvec{v}_1 \in \cefaout(\cB_1), \cdots, \myvec{v}_n \in \cefaout(\cB_n)$ such that $\varphi[\myvec{v}_1/R_{\cB_{1}}, \cdots, $ $\myvec{v}_n/R_{\cB_{n}}]$ is true.
\end{observation}

Let $\aut = (R, Q, \Sigma, \delta, I, F, \alpha)$ be a CEFA and $\costset_\aut = \{\myvec{v} \mid \exists q, a, q'.\ (q, a, q', \myvec{v}) $ $\in \delta\}$. 
Moreover, let $\cU_\aut = (Q, \costset_\aut, \delta', I, F)$ be an NFA over the alphabet $\costset_\aut$ that is obtained from $\aut$ by dropping the accepting condition and ignoring the characters, that is, $\delta'$ comprises tuples $(q, \myvec{v}, q')$ such that $(q, a, q', \myvec{v}) \in \delta$ for $a \in \Sigma$.
%Then $\cefaout(\aut) = \parikh(\NFA')$.
%

\begin{observation}\label{obs-cefa-nfa}
For each CEFA $\aut = (R, Q, \Sigma, \delta, I, F, \alpha)$, 
%
$$\cefaout(\aut) = \left\{\sum \limits_{\myvec{v} \in \costset_\aut} \eta(\myvec{v}) \myvec{v} \left \vert\ \eta \in \parikh(\cU_\aut) \mbox{ and }  \alpha\left[  \sum \limits_{\myvec{v} \in \costset_\aut} \eta(\myvec{v}) \myvec{v} \big{/}  R \right] \mbox{ is true}\right.\right\}.$$
%
\end{observation}

For $i \in [n]$, let $\alpha_i$ be the accepting condition of $\cB_i$. 
Then from Observation~\ref{obs-cefa-nfa}, we know that the following two conditions are equivalent, 
\begin{itemize}

\item there are \hdh{outputs} $\myvec{v}_1 \in \cefaout(\cB_1), \cdots, \myvec{v}_n \in \cefaout(\cB_n)$ such that \\$\varphi[\myvec{v}_1/R_{\cB_{1}}, \cdots, \myvec{v}_n/R_{\cB_{n}}]$ is true, 
\item 
there are $\eta_1 \in \parikh(\cU_{\cB_1}), \cdots, \eta_n \in \parikh(\cU_{\cB_n})$ such that 
\[
\bigwedge \limits_{i \in [n]} \alpha_i\left[ \sum \limits_{\myvec{v} \in \costset_{\cB_i}} \eta_i(\myvec{v}) \myvec{v} \big{/}R_{\cB_i}  \right] \wedge \varphi \left[\sum \limits_{\myvec{v} \in \costset_{\cB_1}} \eta_1(\myvec{v}) \myvec{v} \big{/}R_{\cB_1}, \cdots, \sum \limits_{\myvec{v} \in \costset_{\cB_n}} \eta_n(\myvec{v}) \myvec{v} \big{/}R_{\cB_n} \right]
\]
is true. 
\end{itemize}

Therefore, to solve the non-emptiness of CEFA in $\Lambda'_{x_1}, \cdots, \Lambda'_{x_n}$ w.r.t. $\varphi$, it is sufficient to compute the existential LIA formulas $\psi_{\cU_{\cB_1}}(\fZ_{\costset_{\cB_1}}), \cdots, $ $\psi_{\cU_{\cB_n}}(\fZ_{\costset_{\cB_n}})$ to represent the Parikh images of $\cU_{\cB_1}$, $\cdots$, $\cU_{\cB_n}$ respectively, where $\fZ_{\costset_{\cB_i}} = \{\anivarz_{i,\myvec{v}} \mid \myvec{v} \in \costset_{\cB_i}\}$ for $i \in [n]$, and solve the satisfiability of the following existential LIA formula
\[
\begin{array}{l}
\bigwedge \limits_{i \in [n]} \left(\psi_{\cU_{\cB_i}}(\fZ_{\costset_{\cB_i}}) \wedge \alpha_i\left[\sum\limits_{\myvec{v} \in \costset_{\cB_i}} \anivarz_{i, \myvec{v}} \myvec{v} \big{/}R_{\cB_i}  \right] \right)\  \wedge \\
\ \ \varphi \left[ \sum\limits_{\myvec{v} \in \costset_{\cB_1}} \anivarz_{1, \myvec{v}} \myvec{v} \big{/}R_{\cB_1}, \cdots, \sum\limits_{\myvec{v} \in \costset_{\cB_n}} \anivarz_{n, \myvec{v}} \myvec{v} \big{/}R_{\cB_n} \right].
\end{array}
\]
Intuitively, the integer variables $\anivarz_{i, \myvec{v}}$ represent the number of occurrences of $\myvec{v}$ in the strings accepted by $\cU_{\cB_i}$.

Because the sizes of the LIA formulas $\psi_{\cU_{\cB_1}}(\fZ_{\costset_{\cB_1}}), \cdots, \psi_{\cU_{\cB_n}}(\fZ_{\costset_{\cB_n}})$ are proportional to the sizes (more precisely, the alphabet size, the number of states and transitions) of NFA $\cU_{\cB_1}, \cdots, \cU_{\cB_n}$, and the satisfiability of existential LIA formulas is NP-complete, it is vital to reduce the sizes of these NFAs to improve the performance.

Since
$\sum \limits_{\myvec{v} \in \costset(\cB_i)} \eta(\myvec{v}) \myvec{v} = \sum \limits_{\myvec{v} \in \costset(\cB_i) \setminus \{\myvec{0}\}} \eta(\myvec{v}) \myvec{v}$ for each $i \in [n]$ and $\eta \in \parikh(\cU_{\cB_i})$, it turns out that 
the $\myvec{0}$-labeled transitions in $\cU_{\cB_i}$ do not contribute to the final output $\sum \limits_{\myvec{v} \in \costset(\cB_i)} \eta(\myvec{v}) \myvec{v}$. Therefore, we can apply the following size-reduction technique for $\cU_{\cB_i}$'s.
%Regardless of alphabet, we take register updates as new alphabet symbols and $\myvec{0}$ as $\epsilon$ symbol. Then we are capable of using standard minimization to reduce the automata size.

\medskip
\noindent
\emph{Step 2 (Reducing automata sizes).} For each $i \in [n]$, we view the transitions $(q, \myvec{0}, q')$ in $\cU_{\cB_i}$ as $\varepsilon$-transitions $(q, \varepsilon, q')$, and remove the $\varepsilon$-transitions from $\cU_{\cB_i}$. Then we determinize and minimize the resulting NFA.   \qed

For $i \in [n]$, let  $\cC_i$ denote the DFA obtained from $\cU_{\cB_i}$ by executing Step 2 and $\costset_{\cC_i} := \costset_{\cB_i} \setminus \{\myvec{0}\}$. From the construction, we know that $\parikh(\cC_i) = \prj_{\costset_{\cC_i}}(\parikh(\cU_{\cB_i}))$ for each $i \in [n]$.
Therefore, we compute LIA formulas from $\cC_i$'s, instead of $\cU_{\cB_i}$'s, to represent the Parikh images.

\medskip
\noindent
\emph{Step 3 (Computing an LIA formula).} For each $i \in [n]$, we compute an existential LIA formula $\psi_{\cC_i}(\fZ_{\costset_{\cC_i}})$ from $\cC_i$ to represent $\parikh(\cC_i)$. Then the $\cefadec$ problem is reduced to the satisfiability of the following LIA formula, 
%
\begin{equation}\label{eq:parikh_image}
\begin{array}{l}
\bigwedge \limits_{i \in [n]} \left(\psi_{\cC_i}(\fZ_{\costset_{\cC_i}}) \wedge \alpha_i\left[\sum\limits_{\myvec{v} \in \costset_{\cC_i}} \anivarz_{i, \myvec{v}} \myvec{v} \big{/}R_{\cB_i}  \right] \right) \ \wedge \\
\ \ \varphi \left[ \sum\limits_{\myvec{v} \in \costset_{\cC_1}} \anivarz_{1, \myvec{v}} \myvec{v} \big{/}R_{\cB_1}, \cdots, \sum\limits_{\myvec{v} \in \costset_{\cC_n}} \anivarz_{n, \myvec{v}} \myvec{v} \big{/}R_{\cB_n} \right].
\end{array}
\end{equation}

%\hdh{The final Parikh images are as LIA formula (\ref{eq:parikh_image}). Subsequently, we can employ established SMT solvers to determine the satisfiability of the formula and derive a solution if it is satisfiable. The integer variable solutions are directly extracted from the partial model of the formula. The string variable solutions are obtained through indirect means, whose detailed methodology is discussed in Section \ref{sec-opt-sol-gen}.}

\medskip
\noindent
\emph{Step 4 (Solving the LIA formula).}
We utilize some off-the-shelf SMT solver (e.g. Z3) to solve the satisfiability of the LIA formula in Equation~(\ref{eq:parikh_image}). If the solver reports ``sat'' and returns a solution $\amodel$, then we go to Step 5. Otherwise, the $\cefadec$ problem has no solution.

\hdh{
\medskip
\noindent
\emph{Step 5 (Constructing a solution).}
%For each $i \in [n]$, 
Let $\amodel$ be a solution of the LIA formula in Equation~(\ref{eq:parikh_image}). We want to construct from $\amodel$ and the CEFAs $\cB_i$ for $i \in [n]$, a solution $\amodel'$ of the $\cefadec$ problem. The construction of $\amodel'$ is obtained by constructing $\amodel'(x_i)$ for each $i \in [n]$. 
Let $i \in [n]$. We describe how to construct $\amodel'(x_i)$. 
The free variables $\fZ_{\costset_{\cC_i}}$ correspond to the number of occurrences of characters in $\costset_{\cC_i}$, i.e. the alphabet of the NFA $\cC_i$. Moreover, in $\cC_i$, the characters are just the integer vectors and the characters in the original CEFAs $\cB_i$ used to construct $\cC_i$ are ignored. 
The construction of $\amodel'(x_i)$ is actually to solve the reachability problems in a finite transition system where 
\begin{itemize}
\item the nodes are of the form $(q, (\nvec_{i, \myvec{v}})_{\myvec{v} \in \costset_{\cC_i}})$, where $\nvec_{i, \myvec{v}}$ denotes the number of the remaining occurrences of the integer vector $\myvec{v}$ in $\cB_i$, 
%
\item the transitions are of the form 
$$((q, (\nvec_{i, \myvec{v}})_{\myvec{v} \in \costset_{\cC_i}}), (q, a, q', \myvec{u}), (q', (\nvec'_{i, \myvec{v}})_{\myvec{v} \in \costset_{\cC_i}}))$$
such that $(q, a, q', \myvec{u})$ is a transition in $\cB_i$, $\nvec'_{i, \myvec{u}} = \nvec_{i, \myvec{u}}-1$, moreover, $\nvec'_{i, \myvec{v}} = \nvec_{i, \myvec{v}}$ for every $\myvec{v} \neq \myvec{u}$. 
\end{itemize}
Moreover, the initial and final state of the reachability problem are set to be $(q_0, (\amodel(\anivarz_{i, \myvec{v}}))_{\myvec{v} \in \costset_{\cC_i}})$ and $(q_f, (0,\cdots, 0))$ for some initial state $q_0$ and final state $q_f$ in $\cB_i$. 
If a path from the initial state to the final state is found, then the path gives us a desired string $\amodel'(x_i)$.
}
%Note that the transition system contains exponentially many nodes. As a result, the searching process takes \emph{doubly exponential time} in the worst case, since there are at most doubly exponentially many simple paths. To alleviate the path-explosion problem, in the sequel, we propose two techniques to reduce the number of transitions of the transition system.

%\qed

\hdh{
\subsection{Complexity Analysis of the $\reclsat$ Problem}
The time complexity of regular expression equivalence with complement has been established as nonelementary~\cite{complementary_complexity_1974}. This directly implies that the $\reclsat$ problem exhibits a nonelementary lower bound. We now turn our attention to analyzing the complexity of the $\reclsat$ problem in the absence of complement operations (no occurrence of $\Bar{e}$ and $e_1 \backslash e_2$).

\begin{lemma}\label{lemma:regex2cefa}
    Register-representable regex without complement operation can be converted to CEFA in EXPSPACE.
\end{lemma}

\begin{proof}
    We prove this lemma through structural induction on regular expression $e$. For the base case—where the regular expression consists of a single character, $\epsilon$, or $\emptyset$—the lemma holds trivially.
    For the inductive cases, assuming the lemma holds for regular expressions $e_1$ and $e_2$:
    When $e = e_1\concat e_2$ or $e = e_1+e_2$, the concatenation or union operations described in Section \ref{sec:automaton} produce a CEFA of size $O(2^{|e_1|} + 2^{|e_2|}) = O(2^{\max(|e_1|, |e_2|)}) = O(2^{|e|})$.
    For $e = e_1\cap e_2$, the intersection operation detailed in Section \ref{sec:automaton} yields a CEFA of size $O(2^{|e_1|}\cdot 2^{|e_2|})=O(2^{|e_1| + |e_2|})=O(2^{|e|})$.
    In the case of $e=e_1^{\{m,n\}}$, our construction in Section \ref{subsec:regex2cefa} avoids direct unfolding of counting operators, resulting in a CEFA of size $O(2^{|e_1|}) = O(2^{|e|})$.
    Finally, for $e=(e_1)$ and $e=e^*$, our application of Thompson's construction~\cite{Thompson1968} produces a CEFA of size $O(2^{|e_1|}) = O(2^{|e|})$.
\end{proof}

\begin{theorem}\label{theorem_pspace}
    If all regexes are register-representable and complement operations are excluded, $\reclsat$ can be solved in EXPSPACE.
\end{theorem}
\begin{proof}
    Our decision procedure for $\reclsat$ with only register-representable regexes can be decomposed into two steps: (1) constructing CEFAs from regular membership and length function, (2) solving $\cefadec$.
    Based on Lemma \ref{lemma:regex2cefa}, our construction yields a CEFA of exponential size for each register-representable regex without complement operation. Furthermore, for each string variable's length function, a CEFA of constant size is constructed in Section \ref{subsec:regex2cefa}.
    Consequently, constructing CEFAs in the decision procedure outlined in Section \ref{subsec:regex2cefa} requires nondeterministic exponential size. After constructing CEFAs, the $\reclsat$ problem is transferred to the $\cefadec$ problem. By Proposition \ref{prop-cefadec}, we can demonstrate that the decision procedure likewise requires nondeterministic exponential size. Therefore, the complexity of the decision procedure for solving the restricted $\reclsat$ problem is NEXPSPACE, which yields EXPSPACE by Savitch's theorem~\cite{complexity-book}.
\end{proof}

\begin{theorem}
    $\reclsat$ can be solved in 3-EXPTIME when complement operations are excluded.
\end{theorem}

\begin{proof}
    Our decision procedure for $\reclsat$ with non-register-representable regexes can be decomposed into three steps: (1) unfolding non-register-representable occurrences of counting operators, (2) constructing CEFAs from regular membership and length function, (3) solving $\cefadec$.
    From the decision procedure in Section \ref{subsec:regex2cefa}, unfolding non-register-representable occurrences of counting operators leads to an exponential growth in expression length. For each non-register-representable regular expression of length $l$, the unfolding process generates a register-representable regular expression of length $O(m^n\cdot l)=O((2^{|l|})^{|l|}\cdot l) = O(2^{|l|})$, where $m$ represents the maximum counting bound and $n$ denotes the maximum depth of nested counting operators. By Theorem \ref{theorem_pspace} and the fact that EXPSPACE $\subseteq$ 2-EXPTIME \cite{complexity-book}, the complexity of the decision procedure for solving the $\reclsat$ problem without complement operations is 3-EXPTIME.
\end{proof}
}